---
title: "Stat380_FinalProject_Hainley_Sepich"
author: "Joseph Sepich and Connor Hainley"
date: "2019-04-30"
output: 
  html_document:
    keep_md: yes
  html_notebook: default
---

# Front matter

```{r message = FALSE, warning = FALSE}
# always clean up R environment
rm(list = ls())

# load all packages here
library(mdsr) # book package of utilities
library(stringr) # utility package for strings
library(tidyr) # tidyverse utilities
library(lubridate) # date utility package
library(data.table) # using fread function
library(rvest) # web scraping package
library(readxl) # read execel files
library(readr) # utilities

# user defined functions

# chooseCols - select columns of data according to regex
# createPopulationTable - create population table based off immigration or not
# random_sim - run run_sim with random parameters for beta, alpha, and chi
# readTable - read immigration table from yearbook webpage
# run_sim - run romer growth model simulation

# data
data(compiledData) # our data of GDP and immigration by GeoName and Industry
```
# Obtaining the data

Links to download data from BEA:

GDP by area:  https://apps.bea.gov/regional/downloadzip.cfm
GDP: https://www.bea.gov/data/gdp/gross-domestic-product
Homeland Security: https://www.dhs.gov/immigration-statistics/yearbook/2017


Codes to know:

* GDP in current dollars SAGDP2
* Compensation of Employees SAGDP4
* Real GDP in chained SAGDP9
* Per capita real GDP SAGDP10

## Compile GDP data into tables based on code

```{r}
# First read in all csv files into single data table
dataPath <- read_file('dataPath.txt')
```
```{r}
files <- list.files(path = dataPath,
                    pattern = "^SAGDP2N__ALL_AREAS_[0-9]*_[0-9]*.csv$")
# SAGDP2
nominalGDP <- fread(file = paste0(dataPath,files[1]), fill = TRUE)
head(nominalGDP)
write.csv(nominalGDP, file = paste0(dataPath,"SAGDP2.csv"))
```

```{r}
files <- list.files(path = dataPath,
                    pattern = "^SAGDP4N__ALL_AREAS_[0-9]*_[0-9]*.csv$")
# SAGDP4
compensationGDP <- fread(file = paste0(dataPath,files[1]), fill = TRUE)
head(compensationGDP)
write.csv(compensationGDP, file = paste0(dataPath,"SAGDP4.csv"))
```

```{r}
files <- list.files(path = dataPath,
                    pattern = "^SAGDP9N__ALL_AREAS_[0-9]*_[0-9]*.csv$")
# SAGDP9
realGDP <- fread(file = paste0(dataPath,files[1]), fill = TRUE)
head(realGDP)
write.csv(realGDP, file = paste0(dataPath,"SAGDP9.csv"))
```

```{r}
files <- list.files(path = dataPath,
                    pattern = "^SAGDP10N__ALL_AREAS_[0-9]*_[0-9]*.csv$")
# SAGDP10
perCapitaGDP <- fread(file = paste0(dataPath,files[1]), fill = TRUE)
head(perCapitaGDP)
write.csv(perCapitaGDP, file = paste0(dataPath,"SAGDP10.csv"))
```
 
## Scraping Immigration statistics 
 
```{r}
# returns immigration table found on the page of a specific url
readTable <- function(url) {
  table <- url %>%
            read_html() %>%
            html_nodes(xpath=paste0('/html/body/div[2]/section/div/div/div[2]/',
                                    'div/section/div/div/div/article/div[1]/',
                                    'div[2]/div/div/table')) %>%
            html_table()
  table <- table[[1]]
}
```

```{r}
# Started putting tables online in 2014, prior they are in csv for download.
# Use 2013, 2004 and scrape 2014-2017
dataPath <- paste0(substr(dataPath, 1, nchar(dataPath)-4),"Immigration/")
```
```{r message = FALSE, warning = FALSE}
# Scrape data from 2014-2017
immigrants <- NA
immigrants <- data.frame(stringsAsFactors = FALSE)
year <- 2014
while (year <= 2017) {
  url <- paste0("https://www.dhs.gov/immigration-statistics/yearbook/",year,
                "/table22")
  if (year == 2014) {
    immigrants <- readTable(url)
  } else {
    immigrants[,as.character(year)] <- readTable(url)[,as.character(year)]
  }
  year <- year + 1
}
# Read in and add data from 8_ - 2013
# Table33
# 2103 table 22
files <- list.files(path = dataPath, pattern = ".(xls|xlsx)$")
data1 <- read_excel(paste0(dataPath,files[1]))
data2 <- read_excel(paste0(dataPath,files[2]))

data1 <- data1[1:(length(data1$`Table 22.`)-3),]
data1 <- data1[3:length(data1$`Table 22.`),]
data1[1,1] <- "State or Territory of Residence"

data2 <- data2 %>%
  filter(!is.na(`TABLE 33.`))
data2 <- data2[1:(length(data2$`TABLE 33.`)-2),]
data2 <- data2[2:length(data2$`TABLE 33.`),]
data2[1,1] <- data1[1,1]

colnames(data1) <- as.character(unlist(data1[1,]))
data1 <- data1[-1, ]
data1 <- data1 %>%
  select(-c("2012","2013"))

colnames(data2) <- as.character(unlist(data2[1,]))
data2 <- data2[-1, ]
data2 <- data2 %>%
  select(-c("2004"))

# Convert data types to numeric to match with data1
colsNum <- colnames(data2)
colsNum <- colsNum[2:length(colsNum)]
data2[colsNum] <- sapply(data2[colsNum],as.numeric)
colsNum <- colnames(immigrants)
colsNum <- colsNum[2:length(colsNum)]
immigrants[colsNum] <- sapply(immigrants[colsNum],function(x) gsub(",","",x))
immigrants[colsNum] <- sapply(immigrants[colsNum],as.numeric)

immigrants <- immigrants %>%
      left_join(data1, by = c("State or Territory of Residence"))
immigrants <- immigrants %>%
      left_join(data2, by = c("State or Territory of Residence"))
write.csv(immigrants, file = paste0(dataPath,"immigration.csv"))
```
 
## Read in tables
```{r eval=FALSE}
immigrants <- fread(file = paste0(dataPath,"immigration.csv"))
dataPath <- paste0(substr(dataPath, 1, nchar(dataPath)-12),"GDP/")
nominalGDP <- fread(file = paste0(dataPath,"SAGDP2.csv"))
compensationGDP <- fread(file = paste0(dataPath,"SAGDP4.csv"))
realGDP <- fread(file = paste0(dataPath,"SAGDP9.csv"))
perCapitaGDP <- fread(file = paste0(dataPath,"SAGDP10.csv"))
```

## Join two data sets

Now let us join the GDP data and imigration data into a single table.

```{r warning = FALSE}
econData <- nominalGDP %>%
  rbind(compensationGDP) %>%
  rbind(realGDP) %>%
  rbind(perCapitaGDP) %>%
  filter(nchar(TableName) != 0)
# Name GDP years
names <- colnames(econData)
for (i in 1:length(names)) {
  if(!is.na(as.numeric(names[i]))) {
    econData[,paste0(names[i])] <- as.numeric(econData[,paste0(names[i])])
    names[i] <- paste0("GDP: ",names[i])
  }
}
colnames(econData) <- names
# Name immigation years
names <- colnames(immigrants)
for (i in 1:length(names)) {
  if(!is.na(as.numeric(names[i]))) {
    names[i] <- paste0("Immigrants: ",names[i])
  }
}
colnames(immigrants) <- names
# Change total name
immigrants[1,1] <- "United States"
compiledData <- econData %>%
  left_join(immigrants, by = c("GeoName" = "State or Territory of Residence"))
```



# Simulating the Romer Growth Model

The Romer Growth model is a growth model in economics thats endogenizes the TFP
(Total Factor Productivity) variable. It basically is a production model that
divides labor between production of goods and services and the produciton of
ideas. Below we use the idea that the growth rate of per worker production (per
capita GDP) is based off of beta, exponent of capital in production function, alpha,
percentage of population devoted to production of ideas, and chi, the productivity of
the workers producing ideas. It is also based on the total population.

As an equation we can then state if in steady state:
\[g_{y^*} = \frac{1}{1-\beta}g_A = \frac{1}{1-\beta}\chi\alpha\overline{N}\]

The steady state is the point at which investment in the economy merely replaces
depreciating capital and buying capital for new workers.

## User Functions

Let's create a function to choose columns we want: (like select clause of SQL)
```{r}
chooseCols <- function(x, regex) {
  cols <- grep(regex, names(x), value=TRUE)
  return(x[, cols])
}
```

Next let us calculate the population of the regions based off real GDP and per capita
GDP.

```{r}
# Returns table of population in millions
createPopulationTable <- function(immigrants = TRUE) {
  # Divide real GDP by percapita
  real <- compiledData %>%
    filter(TableName == "SAGDP9N" & IndustryId == 1)
  per <- compiledData %>%
    filter(TableName == "SAGDP10N")
  real <- chooseCols(real, "^(GDP:|GeoName)")
  per <- chooseCols(per, "^(GDP:|GeoName)")
  pop <- real
  cols <- names(per)
  for (i in 2:length(cols)) {
    if(immigrants == TRUE) {
      pop[,paste0(cols[i])] <- real[,paste0(cols[i])] / per[,paste0(cols[i])]
    } else {
      year <- 1995 + i
      immigrants <- compiledData %>%
          filter(TableName == "SAGDP10N") %>%
          chooseCols(paste0("^(Immigrants: ",as.character(year),"|GeoName)"))
      pop[,paste0(cols[i])] <- (real[,paste0(cols[i])] / per[,paste0(cols[i])]) - (immigrants[,2] / 1000000)
    }
  }
  return(pop)
}
```
```{r warning = FALSE}
bop <- createPopulationTable(FALSE)
head(bop)
bop <- createPopulationTable(TRUE)
head(bop)
```

## Getting the data

```{r}
# per capita GDP is tableName SAGDP10N
perCapitaGDP <- compiledData %>%
  filter(TableName == "SAGDP10N")
outputPerWorker <- chooseCols(perCapitaGDP, "^(GDP:|GeoName)")
head(outputPerWorker)
```

## Creating a basic simulation

Now we can create a simulation to test what steady state growth would be with and
without immigrants.

```{r}
# Romer growth parameters
b <- 0.3
a <- 0.2
c <- 0.04
# Population in millions
run_sim <- function(beta, alpha, chi, area = "United States",immigrants = TRUE) {
  row <- grep(area, unique(perCapitaGDP$GeoName))
  population <- createPopulationTable(immigrants)
  growth <- population
  cols <- names(population)
  colsGrowth <- gsub("GDP:","Growth:",cols)
  for (i in 2:length(cols)) {
    growth[,paste0(colsGrowth[i])] <- (1 / (1 - beta)) * alpha * chi * population[,paste0(cols[i])]
  }
  growth <- chooseCols(growth, "^(Growth:|GeoName)")
  # Show graph of PerCapita and growth
  years <- grep("[0-9]$", names(outputPerWorker), value=TRUE)
  years <- as.numeric(gsub("GDP: ","",years))
  totalper <- as.numeric((as.vector(outputPerWorker[row,c(2:22)])))
  data <- data.frame(year = years, output = totalper)
  growthPaths <- data %>%
    ggplot(aes(x = year, y = output)) +
    geom_point()
  
  for (i in 1:length(years)) {
    currentYear <- years[i]
    if (i == 1) {
      yearVal <- seq(from = currentYear, to = (currentYear + 1), by = 0.01)
      delta <- outputPerWorker[row,(i + 1)] * (growth[row,(i + 1)] / 100)
      out <- (yearVal - currentYear) * delta  + outputPerWorker[row,(i + 1)]
    } else {
      currentYears <- seq(from = currentYear, to = (currentYear + 1), by = 0.01)
      yearVal <- c(yearVal, currentYears)
      delta <- outputPerWorker[row,(i + 1)] * (growth[row,(i + 1)] / 100)
      out <- c(out, (currentYears - currentYear) * delta + outputPerWorker[row,(i + 1)])
    }
  }
  slope <- data.frame(yearVal = yearVal, out = out)
  growthPaths <- growthPaths +
    geom_line(data = slope, aes(x = yearVal, y = out, color = "Romer Steady State"))
  growthPaths <- growthPaths +
    ggtitle("GDP Growth over time") +
    xlab("Year") +
    ylab("Per Capita GDP ($)")
  
  avgGrowth <- mean(as.numeric((as.vector(growth[row,c(2:22)]))))
  return(list(growthPaths, avgGrowth))
}
```


When we run the sim with preset values we can see the growth of the economy as
a whole versus the growth of Alaska's economy. When the line connects the dots
the economy is approximately at a steady state, but when the dot is more of a step
from the line, there must be some sort of shock to the economy. Knowing this you can
see that Alaska's economy only changes due to shock's in TFP. This follows assumptions
of the Solow growth model and implies that Alaska is not working to produce it's own
techonologies, which can be reflective of smaller nations around the world.

```{r warning = FALSE}
run_sim(b, a, c, area = "United States", immigrants = TRUE)[[1]]
run_sim(b, a, c, area = "Alaska", immigrants = FALSE)[[1]]
```

## Randomizing the simulation

Now let us randomize the parameters to our growth rate and see how immigrants can effect the
growth rate.

```{r warning = FALSE}
random_sim <- function(area = "United States", immigrants = TRUE) {
  beta <- runif(1)
  alpha <- runif(1)
  chi <- 1 / rexp(1, rate = 0.005) # weight productivity low
  if (chi >= 1) {
    chi <- 1
  }
  run_sim(beta = beta, alpha = alpha, chi = chi, area = area, immigrants = immigrants)
}
simulations <- mosaic::do(40) * random_sim(area = "United States", immigrants = FALSE)
data <- data.frame(growths = unlist(simulations$V2))
data <- data %>%
  filter(growths <= 25)
data %>%
  ggplot(aes(x = growths)) +
  geom_density() +
  xlab("Avg Growth of y") +
  ggtitle("Per Worker Production growth rate (%) without Immigration") +
  geom_rug(aes(color = "Simulated Observed Rate"))
simulations <- mosaic::do(40) * random_sim(area = "United States", immigrants = TRUE)
data <- data.frame(growths = unlist(simulations$V2))
data <- data %>%
  filter(growths <= 25)
data %>%
  ggplot(aes(x = growths)) +
  geom_density() +
  xlab("Avg Growth of y") +
  ggtitle("Per Worker Production growth rate (%) with Immigration") +
  geom_rug(aes(color = "Simulated Observed Rate"))
```

You can see that the growth rates with immigrants are slightly higher, but there
is not much of an impact on the economy as a whole. This would imply that our
current influx of immigrants also does not decrease per capita GDP or per worker
production, because the affect is has on per capita GDP would scale with its
effect on the growth rate. What about a smaller area of the economy with a proportionately
larger amount of immigrants? We can look at Texas.

```{r warning = FALSE}
simulations <- mosaic::do(40, parallel = FALSE) * random_sim(area = "Texas", immigrants = FALSE)
data <- data.frame(growths = unlist(simulations$V2))
data <- data %>%
  filter(growths <= 25)
data %>%
  ggplot(aes(x = growths)) +
  geom_density() +
  xlab("Avg Growth of y") +
  ggtitle("Per Worker Production growth rate (%) without Immigration") +
  geom_rug(aes(color = "Simulated Observed Rate"))
simulations <- mosaic::do(40, parallel = FALSE) * random_sim(area = "Texas", immigrants = TRUE)
data <- data.frame(growths = unlist(simulations$V2))
data <- data %>%
  filter(growths <= 25)
data %>%
  ggplot(aes(x = growths)) +
  geom_density() +
  xlab("Avg Growth of y") +
  ggtitle("Per Worker Production growth rate (%) with Immigration") +
  geom_rug(aes(color = "Simulated Observed Rate"))
```

In Texas you can definitely see that immigration would have a larger effect on the
growth rate of the economy, but this also means it would have a larger short term effect
of decreasing real per capita GDP. In layman's terms in the short run Texans are more
likely to earn less money on average, but in the long run the Texan economy is likely
to grow at a faster rate.

## Processing the simulation

```{r warning = FALSE}
simsRun <- c(75,200,350,500)
withParallel <- c(0,0,0,0)
withoutParallel<- c(0,0,0,0)
timeData <- data.frame(simsRun, withoutParallel = withoutParallel, withParallel = withParallel)
for (i in 1:4) {
  ptm <- Sys.time()
  mosaic::do(simsRun[i]) * random_sim(area = "United States", immigrants = FALSE)
  timeData[i,3] <- Sys.time() - ptm
}
for (i in 1:4) {
  ptm <- Sys.time()
  mosaic::do(simsRun[i], parallel = FALSE) * random_sim(area = "United States", immigrants = FALSE)
  timeData[i,2] <- Sys.time() - ptm
}
head(timeData)
```

Here I wanted to show my reasoning for running without being in parallel for the
earlier simulation. Due to some kind of processing overhead, at only 40 simulations,
it is actually faster to run without parallel processes and at some line between 200
and 500 it does actually become faster to run in parallel. To me this implies there
isn't code that does a redudant/similar thing a lot of the times to make the simulation slow.
